APPENDIX

### 2017-03-23 APPENDIX.tex
Added appendix page

\begin{document}

\IfFileExists{main.pdf}{\includepdf{main.pdf}}{}

\title{Script Grammar}
\date{}
\maketitle

\section{Intdoduction}

Script is a constrained stack-based language inspired by Forth.

Script programs (scripts) exist within Bitcoin transactions, which eventually are finalized or irrevocably committed to the global state of the Bitcoin network by their inclusion in a block which exists in the block-chain with the most proof of work expended.

Transactions consist of inputs and outputs, both of which include scripts. A transaction input script \textit{unlocks} bitcoins; a transaction output script \textit{locks} bitcoins. An input references the output of a previous transaction and provides a script which yields $\sigma_0$ when run. In order for the input to successfully unlock the bitcoins locked up in the referenced output, said output's script must yield $\sigma_1 = \sigma[V = valid]$ when run with it's state initialized to the state $\sigma_0$--yielded by the unlocking/input/spending script.

Script program syntax is defined in terms of byte-commands. Each command is one byte long. Not all bytes are defined as valid commands and many bytes which are defined have syntactic restrictions on the following program bytes which must hold in order for the program to be well-formed. Script defines 173 such commands, each corresponding to a particular byte, also called \textit{words} or \textit{opcodes}. For example, the command that adds 1 to the value on top of the stack, is written \texttt{0x8b}. These opcodes have human-readable aliases to make description of the language easier. For example, the opcode \texttt{0x8b} is referred to as \texttt{OP\_1ADD} in the reference implementation's source code and documentation.

Script is intentionally constrainted: There language does not allow for the expression of loops or recursion. This is desirable, guaranteeing that the runtime of a Script program is linear with the size of the program.

Script is defined by the specification provided by the reference implementation. 
Script is  as sequences of bytecodes. In describing its semantics, we will generally not refer to the format of the values, but some commands in the language only have meaning if we take into account the fact that any part of a Script program is a byte vector. For this reason, we define byte vectors as the following, and will reserve \texttt{b} and \texttt{B} to express a byte and a byte vector, respectively.

\hyperlink{OPSEM}{Section 2} presents a consistent and exhaustive BNF describing Script's syntax.

\hyperlink{OPSEM}{Section 3} specifies the operational semantics of the syntactic constructs defined in \hyperlink{OPSEM}{Section 2}.
\section{Syntax}

\subsection{Fundamental data types}
\noindent
\texttt{b ::= 0x00 | 0x01 | ... | 0xFF} \\
\texttt{B ::= b | b B} \\
\texttt{b$_i$ ::= b} \\
\texttt{B$^n$ ::= b$_1$ b$_2$ ... b$_n$} \\
\texttt{varint ::= }

\subsection{Transactions}
\noindent
\texttt{tx ::= tx\_version tx\_numin tx\_inputs tx\_numout tx\_outputs tx\_locktime  } \\
\texttt{tx\_version ::= B$^4$  } \\
\texttt{tx\_numin ::= varint  } \\
\texttt{tx\_inputs ::= txin | txin tx\_inputs   } \\
\texttt{tx\_numout ::= varint  } \\
\texttt{tx\_outputs ::= txout | txout tx\_outputs  } \\
\texttt{tx\_locktime ::= B$^4$  }

\subsubsection{Transaction Inputs}
\noindent
\texttt{txin ::= txin\_prevhash txin\_index txin\_scriptlen txin\_script txin\_seqno} \\
\texttt{txin\_prevhash ::= B$^{256}$} \\
\texttt{txin\_index ::= B$^4$} \\
\texttt{txin\_scriptlen ::= varint} \\
\texttt{txin\_script ::= script} \\
\texttt{txin\_seqno ::= B$^4$}

\subsubsection{Transaction Outputs}
\noindent
\texttt{txout ::= txout\_value txout\_scriptlen txout\_script  } \\
\texttt{txout\_value ::= B$^8$} \\
\texttt{txout\_scriptlen ::= varint} \\
\texttt{txout\_script ::= script}

\subsection{Script}
\noindent
\texttt{script ::= com | com script} \\
\texttt{com ::= scom | mcom}

\subsubsection{Single-Word Commands}
\noindent \hangindent=0.7cm
\texttt{scom ::=
  scom\_push | scom\_control | scom\_stack | scom\_splice | scom\_bitlogic |
  scom\_numeric | scom\_crypto | scom\_expansion | scom\_template} \\

\noindent \hangindent=0.7cm
\texttt{scom\_push ::=
  OP\_0 | OP\_FALSE | OP\_1NEGATE | \sout{OP\_RESERVED} | OP\_1 | OP\_TRUE |
  OP\_2 | OP\_3 | OP\_4 | OP\_5 | OP\_6 | OP\_7 | OP\_8 | OP\_9 | OP\_10 |
  OP\_11 | OP\_12 | OP\_13 | OP\_14 | OP\_15 | OP\_16} \\

\noindent \hangindent=0.7cm
\texttt{scom\_control ::=
 OP\_NOP | \sout{OP\_VER} | \sout{OP\_VERIF} | \sout{OP\_VERNOTIF} | OP\_VERIFY
 | OP\_RETURN} \\

\noindent \hangindent=0.7cm
\texttt{scom\_stack ::=
  OP\_TOALTSTACK | OP\_FROMALTSTACK | OP\_2DROP | OP\_2DUP | OP\_3DUP |
  OP\_2OVER | OP\_2ROT | OP\_2SWAP | OP\_IFDUP | OP\_DEPTH | OP\_DROP | OP\_DUP
  | OP\_NIP | OP\_OVER | OP\_PICK | OP\_ROLL | OP\_ROT | OP\_SWAP | OP\_TUCK}
  \\

\noindent \hangindent=0.7cm
\texttt{scom\_splice ::=
  \sout{OP\_CAT }| \sout{OP\_SUBSTR} | \sout{OP\_LEFT} | \sout{OP\_RIGHT} |
  OP\_SIZE} \\

\noindent \hangindent=0.7cm
\texttt{scom\_bitlogic ::=
  \sout{OP\_INVERT} | \sout{OP\_AND} | \sout{OP\_OR} | \sout{OP\_XOR} |
  OP\_EQUAL | OP\_EQUALVERIFY | \sout{OP\_RESERVED1} | \sout{OP\_RESERVED2}} \\

\noindent \hangindent=0.7cm
\texttt{scom\_numeric ::=
  OP\_1ADD | OP\_1SUB | \sout{OP\_2MUL} | \sout{OP\_2DIV} | OP\_NEGATE |
  OP\_ABS | OP\_NOT | OP\_0NOTEQUAL | OP\_ADD | OP\_SUB | \sout{OP\_MUL} |
  \sout{OP\_DIV} | \sout{OP\_MOD} | \sout{OP\_LSHIFT} | \sout{OP\_RSHIFT} |
  OP\_BOOLAND | OP\_BOOLOR | OP\_NUMEQUAL | OP\_NUMEQUALVERIFY |
  OP\_NUMNOTEQUAL | OP\_LESSTHAN | OP\_GREATERTHAN | OP\_LESSTHANOREQUAL |
  OP\_GREATERTHANOREQUAL | OP\_MIN | OP\_MAX | OP\_WITHIN} \\

\noindent \hangindent=0.7cm
\texttt{scom\_crypto ::=
  OP\_RIPEMD160 | OP\_SHA1 | OP\_SHA256 | OP\_HASH160 | OP\_HASH256 |
  OP\_CODESEPARATOR | OP\_CHECKSIG | OP\_CHECKSIGVERIFY | OP\_CHECKMULTISIG |
  OP\_CHECKMULTISIGVERIFY} \\

\noindent \hangindent=0.7cm
\texttt{scom\_expansion ::=
  \sout{OP\_NOP1} | OP\_CHECKLOCKTIMEVERIFY | OP\_NOP2 |
  OP\_CHECKSEQUENCEVERIFY | OP\_NOP3 | \sout{OP\_NOP4} | \sout{OP\_NOP5} |
  \sout{OP\_NOP6} | \sout{OP\_NOP7} | \sout{OP\_NOP8} | \sout{OP\_NOP9} |
  \sout{OP\_NOP10}} \\

\noindent \hangindent=0.7cm
\texttt{scom\_template ::=
  \sout{OP\_SMALLINTEGER} | \sout{OP\_PUBKEYS} | \sout{OP\_PUBKEYHASH} |
  \sout{OP\_PUBKEY} | \sout{OP\_INVALIDOPCODE}}

\subsubsection{Multiple-Word Commands}
\noindent \hangindent=0.7cm
\texttt{mcom ::= mcom\_if | mcom\_notif | mcom\_push} \\

% compund control commands
\noindent
\texttt{mcom\_if ::= OP\_IF script OP\_ENDIF | OP\_IF script mcom\_else OP\_ENDIF} \\
\texttt{mcom\_notif ::= OP\_NOTIF script OP\_ENDIF | OP\_NOTIF script mcom\_else OP\_ENDIF} \\
\texttt{mcom\_else ::= OP\_ELSE script | OP\_ELSE script mcom\_else } \\

% compound push commands
\noindent \hangindent=0.7cm
\texttt{mcom\_push ::=
  OP\_PUSHBYTES$_N$ B$^N$ | OP\_PUSHDATA1 B$^1$ B |OP\_PUSHDATA2 B$^2$ B
  | OP\_PUSHDATA4 B$^4$ B}

\subsubsection{Disabled Commands}
\noindent \hangindent=0.7cm
\texttt{dcom ::=
  dcom\_push | dcom\_control | dcom\_stack | dcom\_splice | dcom\_bitlogic |
  dcom\_numeric | dcom\_crypto | dcom\_expansion | dcom\_template} \\

\noindent \hangindent=0.7cm
\texttt{dcom\_push ::= \sout{OP\_RESERVED}} \\

\noindent \hangindent=0.7cm
\texttt{scom\_control ::=
 \sout{OP\_VER} | \sout{OP\_VERIF} | \sout{OP\_VERNOTIF}} \\

\noindent \hangindent=0.7cm
\texttt{dcom\_stack ::= } \\

\noindent \hangindent=0.7cm
\texttt{dcom\_splice ::=
  \sout{OP\_CAT }| \sout{OP\_SUBSTR} | \sout{OP\_LEFT} | \sout{OP\_RIGHT}
  } \\

\noindent \hangindent=0.7cm
\texttt{dcom\_bitlogic ::=
  \sout{OP\_INVERT} | \sout{OP\_AND} | \sout{OP\_OR} | \sout{OP\_XOR} |
  \sout{OP\_RESERVED1} | \sout{OP\_RESERVED2}} \\

\noindent \hangindent=0.7cm
\texttt{dcom\_numeric ::=
  \sout{OP\_2MUL} | \sout{OP\_2DIV} | \sout{OP\_MUL} |
  \sout{OP\_DIV} | \sout{OP\_MOD} | \sout{OP\_LSHIFT} | \sout{OP\_RSHIFT}} \\

\noindent \hangindent=0.7cm
\texttt{dcom\_crypto ::=} \\

\noindent \hangindent=0.7cm
\texttt{dcom\_expansion ::=
  \sout{OP\_NOP1} | \sout{OP\_NOP4} | \sout{OP\_NOP5} |
  \sout{OP\_NOP6} | \sout{OP\_NOP7} | \sout{OP\_NOP8} | \sout{OP\_NOP9} |
  \sout{OP\_NOP10}} \\

\noindent \hangindent=0.7cm
\texttt{dcom\_template ::=
  \sout{OP\_SMALLINTEGER} | \sout{OP\_PUBKEYS} | \sout{OP\_PUBKEYHASH} |
  \sout{OP\_PUBKEY} | \sout{OP\_INVALIDOPCODE}}

\pagebreak

\subsubsection{OP Codes}

\begin{multicols}{3}

\textbf{\centering push value} \\
\texttt{OP\_0 ::= 0x00  } \\
\texttt{OP\_FALSE ::= OP\_0  } \\
\texttt{{OP\_PUSHBYTES$^N$} ::= 0x01-0x4b} \\
\texttt{OP\_PUSHDATA1 ::= 0x4c  } \\
\texttt{OP\_PUSHDATA2 ::= 0x4d  } \\
\texttt{OP\_PUSHDATA4 ::= 0x4e  } \\
\texttt{OP\_1NEGATE ::= 0x4f  } \\
\texttt{OP\_RESERVED ::= 0x50  } \\
\texttt{OP\_1 ::= 0x51  } \\
\texttt{OP\_TRUE::=OP\_1  } \\
\texttt{OP\_2 ::= 0x52  } \\
\texttt{OP\_3 ::= 0x53  } \\
\texttt{OP\_4 ::= 0x54  } \\
\texttt{OP\_5 ::= 0x55  } \\
\texttt{OP\_6 ::= 0x56  } \\
\texttt{OP\_7 ::= 0x57  } \\
\texttt{OP\_8 ::= 0x58  } \\
\texttt{OP\_9 ::= 0x59  } \\
\texttt{OP\_10 ::= 0x5a  } \\
\texttt{OP\_11 ::= 0x5b  } \\
\texttt{OP\_12 ::= 0x5c  } \\
\texttt{OP\_13 ::= 0x5d  } \\
\texttt{OP\_14 ::= 0x5e  } \\
\texttt{OP\_15 ::= 0x5f  } \\
\texttt{OP\_16 ::= 0x60  } \\

\textbf{\centering control} \\
\texttt{OP\_NOP ::= 0x61  } \\
\texttt{OP\_VER ::= 0x62  } \\
\texttt{OP\_IF ::= 0x63  } \\
\texttt{OP\_NOTIF ::= 0x64  } \\
\texttt{OP\_VERIF ::= 0x65  } \\
\texttt{OP\_VERNOTIF ::= 0x66  } \\
\texttt{OP\_ELSE ::= 0x67  } \\
\texttt{OP\_ENDIF ::= 0x68  } \\
\texttt{OP\_VERIFY ::= 0x69  } \\
\texttt{OP\_RETURN ::= 0x6a  } \\

\textbf{\centering stack ops} \\
\texttt{OP\_TOALTSTACK ::= 0x6b  } \\
\texttt{OP\_FROMALTSTACK ::= 0x6c  } \\
\texttt{OP\_2DROP ::= 0x6d  } \\
\texttt{OP\_2DUP ::= 0x6e  } \\
\texttt{OP\_3DUP ::= 0x6f  } \\
\texttt{OP\_2OVER ::= 0x70  } \\
\texttt{OP\_2ROT ::= 0x71  } \\
\texttt{OP\_2SWAP ::= 0x72  } \\
\texttt{OP\_IFDUP ::= 0x73  } \\
\texttt{OP\_DEPTH ::= 0x74  } \\
\texttt{OP\_DROP ::= 0x75  } \\
\texttt{OP\_DUP ::= 0x76  } \\
\texttt{OP\_NIP ::= 0x77  } \\
\texttt{OP\_OVER ::= 0x78  } \\
\texttt{OP\_PICK ::= 0x79  } \\
\texttt{OP\_ROLL ::= 0x7a  } \\
\texttt{OP\_ROT ::= 0x7b  } \\
\texttt{OP\_SWAP ::= 0x7c  } \\
\texttt{OP\_TUCK ::= 0x7d  } \\

\textbf{\centering splice ops} \\
\texttt{OP\_CAT ::= 0x7e  } \\
\texttt{OP\_SUBSTR ::= 0x7f  } \\
\texttt{OP\_LEFT ::= 0x80  } \\
\texttt{OP\_RIGHT ::= 0x81  } \\
\texttt{OP\_SIZE ::= 0x82  } \\

\textbf{\centering bit logic} \\
\texttt{OP\_INVERT ::= 0x83  } \\
\texttt{OP\_AND ::= 0x84  } \\
\texttt{OP\_OR ::= 0x85  } \\
\texttt{OP\_XOR ::= 0x86  } \\
\texttt{OP\_EQUAL ::= 0x87  } \\
\texttt{OP\_EQUALVERIFY ::= 0x88  } \\
\texttt{OP\_RESERVED1 ::= 0x89  } \\
\texttt{OP\_RESERVED2 ::= 0x8a  } \\

\textbf{\centering numeric} \\
\texttt{OP\_1ADD ::= 0x8b  } \\
\texttt{OP\_1SUB ::= 0x8c  } \\
\texttt{OP\_2MUL ::= 0x8d  } \\
\texttt{OP\_2DIV ::= 0x8e  } \\
\texttt{OP\_NEGATE ::= 0x8f  } \\
\texttt{OP\_ABS ::= 0x90  } \\
\texttt{OP\_NOT ::= 0x91  } \\
\texttt{OP\_0NOTEQUAL ::= 0x92  } \\
\texttt{OP\_ADD ::= 0x93  } \\
\texttt{OP\_SUB ::= 0x94  } \\
\texttt{OP\_MUL ::= 0x95  } \\
\texttt{OP\_DIV ::= 0x96  } \\
\texttt{OP\_MOD ::= 0x97  } \\
\texttt{OP\_LSHIFT ::= 0x98  } \\
\texttt{OP\_RSHIFT ::= 0x99  } \\
\texttt{OP\_BOOLAND ::= 0x9a  } \\
\texttt{OP\_BOOLOR ::= 0x9b  } \\
\texttt{OP\_NUMEQUAL ::= 0x9c  } \\
\texttt{OP\_NUMEQUALVERIFY ::= 0x9d  } \\
\texttt{OP\_NUMNOTEQUAL ::= 0x9e  } \\
\texttt{OP\_LESSTHAN ::= 0x9f  } \\
\texttt{OP\_GREATERTHAN ::= 0xa0  } \\
\texttt{OP\_LESSTHANOREQUAL ::= 0xa1  } \\
\texttt{OP\_GREATERTHANOREQUAL ::= 0xa2  } \\
\texttt{OP\_MIN ::= 0xa3  } \\
\texttt{OP\_MAX ::= 0xa4  } \\
\texttt{OP\_WITHIN ::= 0xa5  } \\

\textbf{\centering crypto} \\
\texttt{OP\_RIPEMD160 ::= 0xa6  } \\
\texttt{OP\_SHA1 ::= 0xa7  } \\
\texttt{OP\_SHA256 ::= 0xa8  } \\
\texttt{OP\_HASH160 ::= 0xa9  } \\
\texttt{OP\_HASH256 ::= 0xaa  } \\
\texttt{OP\_CODESEPARATOR ::= 0xab  } \\
\texttt{OP\_CHECKSIG ::= 0xac  } \\
\texttt{OP\_CHECKSIGVERIFY ::= 0xad  } \\
\texttt{OP\_CHECKMULTISIG ::= 0xae  } \\
\texttt{OP\_CHECKMULTISIGVERIFY ::= 0xaf  } \\

\textbf{\centering expansion} \\
\texttt{OP\_NOP1 ::= 0xb0  } \\
\texttt{OP\_CHECKLOCKTIMEVERIFY ::= OP\_NOP2 ::= 0xb1  } \\
\texttt{OP\_CHECKSEQUENCEVERIFY ::= OP\_NOP3 ::= 0xb2  } \\
\texttt{OP\_NOP4 ::= 0xb3  } \\
\texttt{OP\_NOP5 ::= 0xb4  } \\
\texttt{OP\_NOP6 ::= 0xb5  } \\
\texttt{OP\_NOP7 ::= 0xb6  } \\
\texttt{OP\_NOP8 ::= 0xb7  } \\
\texttt{OP\_NOP9 ::= 0xb8  } \\
\texttt{OP\_NOP10 ::= 0xb9  } \\

\textbf{\centering template matching params }\\
\texttt{OP\_SMALLINTEGER ::= 0xfa  } \\
\texttt{OP\_PUBKEYS ::= 0xfb  } \\
\texttt{OP\_PUBKEYHASH ::= 0xfd  } \\
\texttt{OP\_PUBKEY ::= 0xfe  } \\
\texttt{OP\_INVALIDOPCODE ::= 0xff  } \\

\end{multicols}

\pagebreak[2]

\hypertarget{OPSEM}{\section{Operational Semantics -- Big Step Semantics}}

At any given point of a program's execution, the state is entirely described by the following elements:

\begin{itemize}
	\item The stack $S$, whose elements $S_i$ for $1 \leq i \leq |S|$ are indexed starting from the bottom;
    \item The alt-stack $AS$, an auxiliary stack indexed in the same way;
    \item The validity variable $V$, equal to \texttt{true} by default.
\end{itemize}

Any state in which $V = \texttt{false}$ causes the program to immediately terminate with a special exception, handled by the environment in which the program was being run.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Utility rules
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\hypertarget{OPGEN}{\subsection{Generic rules}}
In order to more succinctly define the operational semantics of Script, we define generic rules not tied to any command in the language that express operations common in a stack-based language.

% PUSH
\subsubsection{PUSH}
\inferrule
{
	|S| = L
}
{	
	<\texttt{PUSH } B, \sigma> \Downarrow \sigma[|S| = L+1, S_{|S|} = B]
}
\vspace{3mm}

% DROP
\subsubsection{DROP}
\inferrule
{
	\sigma(|S|) = L \hspace{3mm} 
    B \Downarrow n \hspace{3mm} 
	L \geq n
}
{
	<\texttt{DROP } B, \sigma> \Downarrow \sigma[S_i = S_{i+1} \hspace{3mm} \forall i. \hspace{1mm} n \leq i < L]
}
\vspace{3mm}

% PUSH_ALT
\subsubsection{PUSH\_ALT}
\inferrule
{|AS| = L}
{<\texttt{PUSH\_ALT } B, \sigma> \Downarrow \sigma[|AS| = L+1, AS_{|AS|} = B]}
\vspace{3mm}

% DROP_ALT
\subsubsection{DROP\_ALT}
\inferrule
{
	\sigma(|AS|) = L \hspace{3mm} \sigma(|AS|) \geq B
}
{
	<\texttt{DROP\_ALT } B, \sigma> \Downarrow \sigma[AS_i = AS_{i+1} \hspace{3mm} \forall i. \hspace{1mm} B \leq i < L]
}
\vspace{3mm}

% TEST
\subsubsection{TEST}
\inferrule
{	
	\sigma(|S|) \geq x \hspace{3mm}
	\sigma(S_x) \Downarrow t \hspace{3mm} 
    t = 0
}
{
	<\texttt{TEST $x$}, \sigma> \Downarrow FALSE
}
\vspace{3mm}

\inferrule
{	
	\sigma(|S|) \geq x \hspace{3mm}
	\sigma(S_x) \Downarrow t \hspace{3mm} 
    t \neq 0
}
{
	<\texttt{TEST $x$}, \sigma> \Downarrow TRUE
}
\vspace{3mm}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Constants}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Constants: OP Code Definitions

% OP_0, OP_FALSE
\subsubsection{OP\_0, OP\_FALSE}
Push the byte-vector representing 0 onto the stack. \\

\inferrule
{
	\sigma(|S|) = L \hspace{3mm} 
    B \Downarrow 0
}
{
	<\texttt{OP\_0}, \sigma> \Downarrow \sigma[|S|=L+1, S_{L+1}=B]
}

% OP_N
\subsubsection{OP\_N, \textit{OP\_PUSHNBYTES}}
Push the next $N$ bytes onto the stack. \\

\inferrule
{
	\sigma(|S|) = L
}
{
	<\texttt{OP\_N $b_1 \ldots b_N$}, \sigma> \Downarrow 
	\sigma[|S|=L+1, S_{L+1}=<b_1 \ldots b_N>]
}

% OP_PUSHDATA1
\subsubsection{OP\_PUSHDATA1}
The next byte specifies how many bytes to push onto the stack as a byte-vector. \\

\inferrule
{
    B_1 \Downarrow k \hspace{3mm} 
	\sigma(|S|) = L \hspace{3mm} 
}
{
	<\texttt{OP\_PUSHDATA1 $B_1^1$ $B_2^k$}, \sigma> \Downarrow 
	\sigma[|S|=L+1, S_{L+1}=B_2]
}

%\vspace{3mm}
%\inferrule
%{
%	x = |$B$| \hspace{3mm}
%    b_1 \Downarrow k \hspace{3mm} 
%    x < k 
%}
%{	
%	<\texttt{OP\_PUSHDATA1 $b_1$ $B$ }, \sigma> \Downarrow 
%	\sigma[V = false]
%}

% OP_PUSHDATA2
\subsubsection{OP\_PUSHDATA2}
The next two bytes specify how many bytes to push onto the stack as a byte-vector. \\

\inferrule
{
    B_1 \Downarrow k \hspace{3mm} 
	\sigma(|S|) = L \hspace{3mm} 
}
{
	<\texttt{OP\_PUSHDATA2 $B_1^2$ $B_2^k$}, \sigma> \Downarrow 
	\sigma[|S|=L+1, S_{L+1}=B_2]
}

%\vspace{3mm}
%\inferrule
%{
%	x = |$B$| \hspace{3mm}
%    (b_1 b_2) \Downarrow k \hspace{3mm} 
%    x < k 
%}
%{	
%	<\texttt{OP\_PUSHDATA2 $b_1$ $b_2$ $B$ }, \sigma> \Downarrow 
%	\sigma[V = false]
%}

% OP_PUSHDATA4
\subsubsection{OP\_PUSHDATA4}
The next four bytes specify how many bytes to push onto the stack as a byte-vector. \\

\inferrule
{
    B_1 \Downarrow k \hspace{3mm} 
	\sigma(|S|) = L \hspace{3mm} 
}
{
	<\texttt{OP\_PUSHDATA4 $B_1^4$ $B_2^k$}, \sigma> \Downarrow
	\sigma[|S|=L+1, S_{L+1}=B_2]
}

%\vspace{3mm}
%\inferrule
%{
%	x = |$B$| \hspace{3mm}
%    (b_1 b_2 b_3 b_4) \Downarrow k \hspace{3mm} 
%    x < k 
%}
%{	
%	<\texttt{OP\_PUSHDATA4 $b_1$ $b_2$ $b_3$ $b_4$ $B$ }, \sigma> \Downarrow 
%	\sigma[V = false]
%}

% OP_1NEGATE
\subsubsection{OP\_1NEGATE}
Push the byte-vector representing -1 onto the stack.\\

\inferrule
{
	\sigma(|S|) = L \hspace{3mm} 
    B \Downarrow -1
}
{<\texttt{OP\_0}, \sigma> \Downarrow 
	\sigma[|S|=l+1, S_{l+1}=B]}

% OP_1, OP_TRUE
\subsubsection{OP\_1, OP\_TRUE}
Push the byte-vector representing 1 onto the stack. \\
\inferrule
{\sigma(|S|) = L \hspace{3mm} B \Downarrow 1}
{<\texttt{OP\_0}, \sigma> \Downarrow 
	\sigma[|S|=L+1, S_{L+1}=B]}

% OP_2-OP_16
\subsubsection{OP\_2-OP\_16, \textit{OP\_PUSHN}}
Push the byte-vector representing the number specified in the word name onto the stack. \\

\inferrule
{\sigma(|S|) = L \hspace{3mm} B \Downarrow N}
{<\texttt{OP\_PUSHN}, \sigma> \Downarrow 
	\sigma[|S|=L+1, S_{L+1}=B]}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Flow Control}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Flow Control: OP Code Definitions

% OP_NOP
\subsubsection{OP\_NOP}
Do nothing. \\

\inferrule
{\quad}
{
	<\texttt{OP\_NOP},\sigma> \Downarrow \sigma
}

% OP_IF, OP_ELSE, OP_ENDIF
\subsubsection{OP\_IF, OP\_ELSE, OP\_ENDIF}
Run $C_1$ if top of stack is present and evaluates to true. Otherwise run $C_2$ if stack is present and evaluates to false. \\

\inferrule
{	
	\sigma(|S|)=L \hspace{3mm}
	<\texttt{TEST $L$},\sigma> \Downarrow \texttt{TRUE} \hspace{3mm}
	<\texttt{DROP $L$},\sigma> \Downarrow \sigma_1 \hspace{3mm}
    <C_1, \sigma_1> \Downarrow \sigma '
}
{	
	<\texttt{OP\_IF $C_1$ OP\_ENDIF}> \Downarrow \sigma '
}
\vspace{3mm}

\inferrule
{	
	\sigma(|S|)=L \hspace{3mm}
	<\texttt{TEST $L$},\sigma> \Downarrow \texttt{FALSE} \hspace{3mm}
	<\texttt{DROP $L$},\sigma> \Downarrow \sigma ' \hspace{3mm}
}
{	
	<\texttt{OP\_IF $C_1$ OP\_ENDIF}> \Downarrow \sigma '
}
\vspace{3mm}

\inferrule
{	\sigma(|S|)=L \hspace{3mm}
	<\texttt{TEST $L$},\sigma> \Downarrow \texttt{TRUE} \hspace{3mm}
	<\texttt{DROP $L$},\sigma> \Downarrow \sigma_1 \hspace{3mm}
    <C_1, \sigma_1> \Downarrow \sigma '
}
{	<\texttt{OP\_IF $C_1$ OP\_ELSE $C_2$ OP\_ENDIF}> \Downarrow \sigma '
}
\vspace{3mm}

\inferrule
{	\sigma(|S|)=L \hspace{3mm}
	<\texttt{TEST $L$},\sigma> \Downarrow \texttt{FALSE} \hspace{3mm}
	<\texttt{DROP $L$},\sigma> \Downarrow \sigma_1 \hspace{3mm}
    <C_2, \sigma_1> \Downarrow \sigma '
}
{	<\texttt{OP\_IF $C_1$ OP\_ELSE $C_2$ OP\_ENDIF}> \Downarrow \sigma '
}
\vspace{3mm}

% OP_NOTIF, OP_ELSE, OP_ENDIF
\subsubsection{OP\_NOTIF, OP\_ELSE, OP\_ENDIF}
Run $C_1$ if top of stack is present and evaluates to false. Otherwise run $C_2$ if stack is present and evaluates to true. \\

\inferrule
{	\sigma(|S|)=L \hspace{3mm}
	<\texttt{TEST $L$},\sigma> \Downarrow \texttt{FALSE} \hspace{3mm}
	<\texttt{DROP $L$},\sigma> \Downarrow \sigma_1 \hspace{3mm}
    <C_1, \sigma_1> \Downarrow \sigma '
}
{	<\texttt{OP\_NOTIF $C_1$ OP\_ENDIF}> \Downarrow \sigma '
}
\vspace{3mm}

\inferrule
{	\sigma(|S|)=L \hspace{3mm}
	<\texttt{TEST $L$},\sigma> \Downarrow \texttt{TRUE} \hspace{3mm}
	<\texttt{DROP $L$},\sigma> \Downarrow \sigma ' \hspace{3mm}
}
{	<\texttt{OP\_NOTIF $C_1$ OP\_ENDIF}> \Downarrow \sigma '
}
\vspace{3mm}

\inferrule
{	\sigma(|S|)=L \hspace{3mm}
	<\texttt{TEST $L$},\sigma> \Downarrow \texttt{FALSE} \hspace{3mm}
	<\texttt{DROP $L$},\sigma> \Downarrow \sigma_1 \hspace{3mm}
    <C_1, \sigma_1> \Downarrow \sigma '
}
{	<\texttt{OP\_NOTIF $C_1$ OP\_ELSE $C_2$ OP\_ENDIF}> \Downarrow \sigma '
}
\vspace{3mm}

\inferrule
{	\sigma(|S|)=L \hspace{3mm}
	<\texttt{TEST $L$},\sigma> \Downarrow \texttt{TRUE} \hspace{3mm}
	<\texttt{DROP $L$},\sigma> \Downarrow \sigma_1 \hspace{3mm}
    <C_2, \sigma_1> \Downarrow \sigma '
}
{	<\texttt{OP\_NOTIF $C_1$ OP\_ELSE $C_2$ OP\_ENDIF}> \Downarrow \sigma '
}
\vspace{3mm}

% OP_VERIFY
\subsubsection{OP\_VERIFY := 0x69}
If top of stack is present and evaluates to true then remove top of stock and mark transaction as valid; otherwise mark transaction as invalid. \\

\inferrule
{	\sigma(|S|)=L \hspace{3mm}
	<\texttt{TEST $L$},\sigma> \Downarrow \texttt{TRUE} \hspace{3mm}
	<\texttt{DROP $L$},\sigma> \Downarrow \sigma ' \hspace{3mm}
    \sigma'(V)=\texttt{invalid}
}
{	<\texttt{OP\_VERIFY}, \sigma> \Downarrow \sigma '
}

% OP_RETURN
\subsubsection{OP\_RETURN := 0x6a}
Unconditionally mark transaction as invalid. \\

\inferrule
{	\quad
}
{	<\texttt{OP\_RETURN}, \sigma> \Downarrow \sigma[V=\texttt{invalid}]
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Stack}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% OP_TOALTSTACK
\subsubsection{OP\_TOALTSTACK}

\inferrule
{	
	|S| = L \hspace{3mm}
    S_L = x \hspace{3mm}
    <\texttt{DROP } L, \sigma> \Downarrow \sigma_1
    <\texttt{PUSHALT } x, \sigma_1> \Downarrow \sigma'
}
{
	<\texttt{OP\_TOALTSTACK}, \sigma > \Downarrow \sigma
}
\vspace{3mm}

% OP_FROMALTSTACK
\subsubsection{OP\_FROMALTSTACK}

\inferrule
{	
	|AS| = L \hspace{3mm}
    AS_L = x \hspace{3mm}
    <\texttt{DROPALT } L, \sigma> \Downarrow \sigma_1
    <\texttt{PUSH } x, \sigma_1> \Downarrow \sigma'
}
{
	<\texttt{OP\_FROMALTSTACK}, \sigma > \Downarrow \sigma
}
\vspace{3mm}

%  OP_IFDUP
\subsubsection{OP\_IFDUP}
If top of stack is present and non-zero, duplicate the top of the stack. \\

\inferrule
{
	\sigma(S_{|S|}) = 0
}
{	
	<\texttt{OP\_IFDUP}, \sigma > \Downarrow \sigma
}
\vspace{3mm}

\inferrule
{
	\sigma(S_{|S|}) \neq 0 \hspace{3mm}
    <\texttt{PUSH } S_{|S|} \Downarrow \sigma ' >
} 
{
	<\texttt{OP\_IFDUP}, \sigma > \Downarrow \sigma'
}
\vspace{3mm}

% OP_DEPTH
\subsubsection{OP\_DEPTH}
Push the byte-vector representing the depth of the stack onto the stack. \\

\inferrule
{
	\sigma(|S|) = L \hspace{3mm} 
    < \texttt{PUSH } L, \sigma > \Downarrow \sigma '
}
{
	<\texttt{OP\_DEPTH}, \sigma > \Downarrow \sigma '
}
\vspace{3mm}

% OP_DROP
\subsubsection{OP\_DROP}
Remove the top of the stack. \\

>>>>

\pagenumbering{gobble}

\begin{document}
\title{APPENDIX}
\date{}
\maketitle
\end{document}
